---
subtitle:    Modelling subtitle
description: |
  - model selection
number-sections: false
toc: true         # show page‐TOC (right‐sidebar) on _every_ page
toc-depth: 3      # include ## and ### in the TOC
---

# Modelling {#sec-14-modelling}

```{r}
#| label: setup
source("../_book_functions.R")
```

### Selection of Modelling Techniques {#sec-14-techniques}

<!-- Outline the statistical and machine learning techniques considered (e.g., ETS, ARIMA, Prophet, Hierarchical Time Series). -->

> The selection of modelling techniques must balance interpretability, accuracy, data availability, and computational cost. Given *Pythia’s Advice* aims to build trust and provide transparent, value-added forecasts with minimal resources, **statistical models** were deliberately chosen for their clarity and theoretical grounding. While this choice limits the ability to model promotions and other exogenous factors — potentially reducing accuracy — it supports stakeholder confidence and facilitates adoption. In a future phase, once trust is established, **machine learning** techniques may be explored to further improve forecast performance.


6.1. Literature Review / Similar Models - comprehensive overview of the state of the art in machine learning, including the latest algorithms, - insight into the performance of different algorithms and techniques on similar types of data. (avoid wasting time on models that are unlikely to perform well.)

---

### Data Preparation for Modelling {#sec-14-data-prep}

Describe how the data was structured for modelling, including train/test split, resampling, lag features, and hierarchical structure.

---

### Modelling Assumptions {#sec-14-assumptions}

List any modelling assumptions (e.g., forecast granularity, univariate/multivariate inputs, independence between nodes).

---

### Model Training and Calibration {#sec-14-training}

Explain the training approach: cross-validation strategy, hyperparameter tuning, recursive vs. direct forecasting.

---

### Forecast Generation {#sec-14-generation}

Describe how forecasts are generated for each technique, including aggregation levels and forecast horizons.

---

### Model Validation and Performance Metrics {#sec-14-validation}

Present the evaluation methodology, error metrics (e.g., RMSE, MAPE, bias), and visualizations.

---

### Model Interpretability and Explainability {#sec-14-explainability}

(Optional) Include techniques for interpreting the forecasts (e.g., SHAP values, residual analysis).

---

### Tools and Frameworks Used {#sec-14-tools}

<!-- List packages, libraries, or software used for modelling (e.g., `fable`, `forecast`, `lightgbm`, DuckDB). -->

#### R & R-Studio tooling

```{r}
#| label: r-version
#| 
data.table(R.Version = R.Version()) %>%
  kbl(
    col.names = names(.),
    align     = "l",
    booktabs  = TRUE,
    escape = FALSE
  ) %>%
  kable_styling(
    full_width          = FALSE          ,
    latex_options = c("hold_position"),
    position            = "left",
    bootstrap_options   = c("striped", "hover", "condensed")
  )

data.table(RStudio.Version = rstudioapi::versionInfo()) %>%
  kbl(
    col.names = names(.),
    align     = "l",
    booktabs  = TRUE,
    escape = FALSE
  ) %>%
  kable_styling(
    full_width          = FALSE          ,
    latex_options = c("hold_position"),
    position            = "left",
    bootstrap_options   = c("striped", "hover", "condensed")
  )
```

#### R-Packages

::: {#tbl-14-packages .table}

```{r}
library(data.table)
library(magrittr)
library(kableExtra)

fread(text = "
Package        , Purpose                                  , Citation
padt           , Pythia’s Advice – Data Tooling           , @padt2024
scales         , Scaling and formatting                   , @scales2024
collapsibleTree, Collapsible trees                        , @collapsibleTree2024
crayon         , Colored text output                      , @crayon2024
treemapify     , Treemaps                                 , @treemapify2024
skimr          , Data summarization                       , @skimr2024
quarto         , Rendering the report                     , @quarto2024
duckDB         , DuckDB database and light transformations, @duckdb2024
gt             , Creating tables                          , @gt2024
DT             , Interactive tables                       , @DT2024
knitr          , Dynamic report generation                , @knitr2024
kableExtra     , Enhanced tables                          , @kableExtra2024
ggrepel        , Non-overlapping labels in ggplot2        , @ggrepel2024
openxlsx2      , Read/write Excel                         , @openxlsx2_2024
fpp3           , Forecasting and time series              , @fpp3_2024
feasts         , Time series feature extraction           , @feasts_2024
tsibble        , Tidy temporal data frames                , @tsibble2024
magrittr       , Pipe operator                            , @magrittr2024
glue           , String interpolation                     , @glue2024
tidyr          , Data tidying                             , @tidyr2024
tidyverse      , Unified data science tools               , @tidyverse2024
purrr          , Functional programming                   , @purrr2024
lubridate      , Date-time manipulation                   , @lubridate2024
ggplot2        , Data visualization                       , @ggplot2_2024
data.table     , Fast data manipulation                   , @data.table2024
dplyr          , Data manipulation                        , @dplyr2024
fabletools     , Forecasting infrastructure               , @fabletools2024
fable          , Forecasting models                       , @fable2024
") %>%
    kbl(
,
    # format = "latex"  ,      
    col.names           = names(.),
    align               = "l",
    booktabs            = TRUE,
    escape = FALSE
  ) %>%
  kable_styling(
    full_width          = FALSE          ,
    latex_options = c("hold_position"),
    position            = "left",
    bootstrap_options   = c("striped", "hover", "condensed")
  )


```

:::



---

### Summary and Lessons Learned {#sec-14-summary}

Summarize findings from the modelling phase and any key takeaways to inform the evaluation phase.

---



### Model Selection

different models have different strengths and weaknesses and are suitable for different types of data and problems. achieve the best possible performance and maximize the impact

avoid overfitting a trade-off between the simplicity and flexibility (robustness)

model performance - only article hierarchy - article and customer hierarchy

Cross-validation is a method for evaluating the performance of a model on a validation dataset. This involves dividing the data into k folds, training the model on k-1 folds, and evaluating the performance on the remaining fold. This process is repeated k times, with each fold serving as the validation set once, and the results are averaged to obtain a final performance score. This method provides a more reliable estimate of the model's performance, as it uses all of the data for training and evaluation.

algorithm - handle teh specific characteristics of teh data - balance performance and interpretability - memory and processing power

-   validity -\> fa & prediction intervals
-   robustness -\>
-   transparency -\>

| Capability                   | Standard ARIMA                                        | SARIMA                         | ARIMAX                                        |
| ---------------------------- | ----------------------------------------------------- | ------------------------------ | --------------------------------------------- |
| Handles Trend                | ✅ (via differencing)                                  | ✅ (with seasonal differencing) | ✅                                             |
| Damped Trend                 | ❌ (not explicit)                                      | ❌                              | ❌                                             |
| Seasonality                  | ❌ (needs SARIMA)                                      | ✅                              | ✅                                             |
| Promotions / External Inputs | ❌                                                     | ❌                              | ✅                                             |
| Best For                     | Stationary or differenced series with autocorrelation | Seasonal time series           | Time series influenced by external regressors |



<!-- #### Modeling Technique -->

<!-- #### Modeling Assumptions -->

<!-- ### Generate Test Design -->

<!-- #### Test Design -->

<!-- #### Build Model -->

<!-- ### Parameter Settings -->

<!-- #### Models -->

<!-- #### Model Description -->

<!-- ### Assess Model -->

<!-- #### Model Assessment -->

<!-- #### Revised Parameter Settings -->


# Appendix {.unnumbered}

<!-- {{< include 04_modelling_def.qmd >}} -->
