---
subtitle:    Modelling subtitle
description: |
  - model selection
number-sections: false
toc: true         # show page‐TOC (right‐sidebar) on _every_ page
toc-depth: 3      # include ## and ### in the TOC
---

# Modelling {#sec-14-modelling}

```{r}
#| label: setup
source("../_book_functions.R")
```

### Selection of Modelling Techniques {#sec-14-techniques}

## Literature {#sec-literature_review}

The literature review focuses on the intersection of three fields, according to the now-ubiquitous Data Science Venn Diagram of [Drew Conway](http://drewconway.com/zia/2013/3/26/the-data-science-venn-diagram).

## Business Process

-   SCOR MODEL A COMPLETE GUIDE - 2020 EDITION, [@blokdykSCORMODELCOMPLETE2019]
-   Inventory and production management in supply chains, [@silverInventoryProductionManagement2021]

## Math & Statistics

-   Forecasting: Principles and Practice (3rd ed), [@hyndmanForecastingPrinciplesPractice2021]
-   Demand forecasting best practices, [@vandeputDemandForecastingBest2023]
-   Data science for supply chain forecasting, [@vandeputDataScienceSupply2021]
-   Inventory optimization: models and simulations, [@vandeputInventoryOptimizationModels2020]
-   Machine Learning With Boosting: A Beginner's Guide, [@hartshornMachineLearningBoosting2017]
-   Introduction to conformal prediction with Python, [@molnarIntroductionConformalPrediction2023]
-   Practical guide to applied conformal prediction in Python, [@manokhinPracticalGuideApplied2023]
-   Mastering Modern Time Series Forecasting : The Complete Guide to Statistical, Machine Learning & Deep Learning Models in Python [@valeriymanokhinMasteringModernTime2025]

<!-- Outline the statistical and machine learning techniques considered (e.g., ETS, ARIMA, Prophet, Hierarchical Time Series). -->

> The selection of modelling techniques must balance interpretability, accuracy, data availability, and computational cost. Given *Pythia’s Advice* aims to build trust and provide transparent, value-added forecasts with minimal resources, **statistical models** were deliberately chosen for their clarity and theoretical grounding. While this choice limits the ability to model promotions and other exogenous factors — potentially reducing accuracy — it supports stakeholder confidence and facilitates adoption. In a future phase, once trust is established, **machine learning** techniques may be explored to further improve forecast performance.

6.1. Literature Review / Similar Models - comprehensive overview of the state of the art in machine learning, including the latest algorithms, - insight into the performance of different algorithms and techniques on similar types of data. (avoid wasting time on models that are unlikely to perform well.)

------------------------------------------------------------------------

### Data Preparation for Modelling {#sec-14-data-prep}

Describe how the data was structured for modelling, including train/test split, resampling, lag features, and hierarchical structure.

------------------------------------------------------------------------

### Modelling Assumptions {#sec-14-assumptions}

List any modelling assumptions (e.g., forecast granularity, univariate/multivariate inputs, independence between nodes).

------------------------------------------------------------------------

### Model Training and Calibration {#sec-14-training}

Explain the training approach: cross-validation strategy, hyperparameter tuning, recursive vs. direct forecasting.

------------------------------------------------------------------------

### Forecast Generation {#sec-14-generation}

Describe how forecasts are generated for each technique, including aggregation levels and forecast horizons.

------------------------------------------------------------------------

### Model Validation and Performance Metrics {#sec-14-validation}

Present the evaluation methodology, error metrics (e.g., RMSE, MAPE, bias), and visualizations.

------------------------------------------------------------------------

### Model Interpretability and Explainability {#sec-14-explainability}

(Optional) Include techniques for interpreting the forecasts (e.g., SHAP values, residual analysis).

------------------------------------------------------------------------

### Tools and Frameworks Used {#sec-14-tools}

<!-- List packages, libraries, or software used for modelling (e.g., `fable`, `forecast`, `lightgbm`, DuckDB). -->

#### R & R-Studio tooling

see [@rcoreteamLanguageEnvironmentStatistical2024] & [@Posit]

```{r}
#| label: r-version
#| 
 # data.table(R.Version = R.Version()) %>%
 #   kbl(
 #     col.names = names(.),
 #     align     = "l",
 #     booktabs  = TRUE,
 #     escape = FALSE
 #   ) %>%
 #   kable_styling(
 #     full_width          = FALSE          ,
 #     latex_options = c("hold_position"),
 #     position            = "left",
 #     bootstrap_options   = c("striped", "hover", "condensed")
 #   )
 
info <- sessionInfo()

data.table(
  Component = c("R Version", "Platform", "Running OS"),
  Value = c(
    info$R.version$version.string,
    info$R.version$platform,
    info$running
  )
) %>%
  kbl(
    booktabs = TRUE,
    escape = FALSE,
    align = "ll"
  ) %>%
  kable_styling(
    full_width = FALSE,
    position = "left",
    bootstrap_options = c("striped", "hover", "condensed"),
    latex_options = c("hold_position")
  )
```

#### R-Packages

::: {#tbl-14-packages .table-wrapper style="text-align: left;"}
| Package | Purpose | Citation |
|:-----------------|:-----------------------------------|:-----------------|
| collapsibleTree | Collapsible trees | @liakhovitskiCollapsibleTreeInteractiveCollapsible2024 |
| crayon | Colored text output | @csardiCrayonColoredTerminal2024 |
| data.table | Fast data manipulation | @barrettDatatableExtensionDataframe2024 |
| dplyr | Data manipulation | @wickhamDplyrGrammarData2024 |
| DT | Interactive tables | @xieDTWrapperJavaScript2024 |
| duckDB | DuckDB database and light transformations | @raasveldtDuckdbInProcessSQL2024 |
| fable | Forecasting models | @ohara-wildFableForecastingModels2024 |
| fabletools | Forecasting infrastructure | @ohara-wildFabletoolsCoreTools2024 |
| feasts | Time series feature extraction | @ohara-wildFeastsFeatureExtraction2024 |
| fpp3 | Forecasting and time series | @hyndmanFpp3ForecastingPrinciples2024 |
| ggplot2 | Data visualization | @wickhamGgplot2CreateElegant2024 |
| ggrepel | Non-overlapping labels in ggplot2 | @slowikowskiGgrepelAutomaticallyPosition2024 |
| glue | String interpolation | @hesterGlueInterpretedString2024 |
| gt | Creating tables | @iannoneGtEasilyCreate2024 |
| kableExtra | Enhanced tables | @zhuKableExtraConstructComplex2024 |
| knitr | Dynamic report generation | @xieKnitrGeneralPurposePackage2024 |
| lubridate | Date-time manipulation | @spinuLubridateMakeDealing2024 |
| magrittr | Pipe operator | @magrittrMagrittrForwardPipeOperator2022 |
| openxlsx2 | Read/write Excel | @barboneOpenxlsx2ReadWrite2025 |
| **padt** | **Pythia’s Advice – Data Tooling** | @EAISIPADT |
| purrr | Functional programming | @wickhamPurrrFunctionalProgramming2025 |
| quarto | Rendering the report | @allaireQuartoInterfaceQuarto2024 |
| scales | Scaling and formatting | @wickhamScalesScaleFunctions2025 |
| skimr | Data summarization | @waringSkimrCompactFlexible2022 |
| tidyr | Data tidying | @wickhamTidyrTidyMessy2024 |
| tidyverse | Unified data science tools | @wickhamTidyverseEasilyInstall2024 |
| treemapify | Treemaps | @wilkinsTreemapifyDrawTreemaps2024 |
| tsibble | Tidy temporal data frames | @wangTsibbleTidyTemporal2024 |
:::

------------------------------------------------------------------------

### Summary and Lessons Learned {#sec-14-summary}

Summarize findings from the modelling phase and any key takeaways to inform the evaluation phase.

------------------------------------------------------------------------

### Model Selection

different models have different strengths and weaknesses and are suitable for different types of data and problems. achieve the best possible performance and maximize the impact

avoid overfitting a trade-off between the simplicity and flexibility (robustness)

model performance - only article hierarchy - article and customer hierarchy

Cross-validation is a method for evaluating the performance of a model on a validation dataset. This involves dividing the data into k folds, training the model on k-1 folds, and evaluating the performance on the remaining fold. This process is repeated k times, with each fold serving as the validation set once, and the results are averaged to obtain a final performance score. This method provides a more reliable estimate of the model's performance, as it uses all of the data for training and evaluation.

algorithm - handle teh specific characteristics of teh data - balance performance and interpretability - memory and processing power

-   validity -\> fa & prediction intervals
-   robustness -\>
-   transparency -\>

| Capability | Standard ARIMA | SARIMA | ARIMAX |
|------------------|-------------------|------------------|------------------|
| Handles Trend | ✅ (via differencing) | ✅ (with seasonal differencing) | ✅ |
| Damped Trend | ❌ (not explicit) | ❌ | ❌ |
| Seasonality | ❌ (needs SARIMA) | ✅ | ✅ |
| Promotions / External Inputs | ❌ | ❌ | ✅ |
| Best For | Stationary or differenced series with autocorrelation | Seasonal time series | Time series influenced by external regressors |

<!-- #### Modeling Technique -->

<!-- #### Modeling Assumptions -->

<!-- ### Generate Test Design -->

<!-- #### Test Design -->

<!-- #### Build Model -->

<!-- ### Parameter Settings -->

<!-- #### Models -->

<!-- #### Model Description -->

<!-- ### Assess Model -->

<!-- #### Model Assessment -->

<!-- #### Revised Parameter Settings -->

# Appendix {.unnumbered}

<!-- {{< include 04_modelling_def.qmd >}} -->
