---
subtitle:     what data is available and where it comes from 
description: |
  - SAP BI extracts serve as the foundation: article, customer, sales & forecast data
  - Data was scoped to finished goods under Make-to-Stock with ambient storage
  - The modeling dataset is based on the top-performing MSO and A-class articles
  - Exploratory analysis confirmed temporal seasonality and product hierarchy depth
  - Data quality is high overall, with minimal missing values and no major outliers
number-sections: false
toc: true         # show page‐TOC (right‐sidebar) on _every_ page
toc-depth: 3      # include ## and ### in the TOC
---

# Data Understanding {#sec-02-data-understanding}

```{r}
#| label: setup
source("../_book_functions.R")
```

## Identification of Data Sources {#sec-02-identification}

<!-- Describe what data is available and where it comes from. -->
Table @tbl-02-data-sources summarizes the data sources used in this project, 
detailing their type, time span, granularity, system of record (SoR), and content. 
All data is sourced from SAP BI, which extracts it from the underlying 
systems of record. 
The table also lists the specific SAP tables used for each data source.  

::: {#tbl-02-data-sources}

```{r}

# Read the data into a data.table 
dtDS <-  fread(text = "
  Data Source, Type, Time-Span, Granularity, SoR   , Table, Content
  article    , MD  ,          ,            , SAP S4, MARC , article attr.  & hierachy
  customer   , MD  ,          ,            , SAP S4, KNVV , customer attr. hierachy
  sales      , TD  , > 2020   , daily      , SAP S4, LIPS , customer orders + free-of-charge 
  forecasts  , TD  , > 2023   , monhtly    , QAD   ,      , monthly pre- & post Demand review forecasts
")

# Create a styled kableExtra table
dtDS %>% 
  kbl(
    # caption             = "Overview of data sources used in the project",
    col.names           = names(.),
    align               = "l",
    booktabs            = TRUE
  ) %>% 
  kable_styling(
    full_width          = FALSE,
    position            = "left",
    bootstrap_options   = c("striped", "hover", "condensed")
  )  
```
Overview of data sources used in the project  
MD = Master Data, TD = Transaction Data
:::

<!-- PDF -->
<!-- %>% -->
<!--   footnote( -->
<!--     general           = "MD = Master data, TD = Transaction Data", -->
<!--     footnote_as_chunk = TRUE, -->
<!--     threeparttable    = FALSE, -->
<!--     escape            = FALSE -->
<!--   ) -->

<!-- Subsections: -->

### Articles {#sec-02-articles}

The Article Master Data is extracted from `MARC` which contains Plant specific 
data, this is enhanced with sales organization specific data from `MVKE` and 
global article data from `MARA`, applied filters are shown in Table @tbl-02-data-sources-article-filters.

::: {#tbl-02-data-sources-article-filters}
```{r}

dtAF <-  fread(text = "
  FIELD      , Filter                        , Reason
  MARA-MTART , HAWA; FERT; VOLL              , only finished goods
  MARA-STRGRP, Z1                            , only Make-to-Stock   
  MARC-DISMM , ZC; ZD; ZG; ZH; ZI; ZJ; ZN; ZP, only MRP relevant
  MARC-PERKZ , W                             , only Ambient
  ")

# Create a styled kableExtra table
dtAF %>% 
  kbl(
    caption             = "Article Master Data Filters",
    col.names           = names(.),
    align               = "l",
    booktabs            = TRUE
  ) %>% 
  kable_styling(
    full_width          = FALSE,
    position            = "left",
    bootstrap_options   = c("striped", "hover", "condensed"),
    latex_options       = c("striped", "hold_position")    
  )
```

Article Master Data Filters  
see @sec-01-bu-scope for reasons for article-filters.
:::

### Customers {#sec-02-customers}

Customer Master Data is sourced from `KNVV`, which contains sales-organization-specific information, and is enriched with global attributes from `KNA1`. This includes customer classification fields such as type (e.g., internal, external, one-time customer, or employee). Customers flagged as employees are excluded from the analysis.

### Sales {#sec-02-sales}
<!-- Data filters applied (Doctype's, Item category, ...) -->
<!-- exclude Promotion articles --> 

Sales data is extracted from `LIPS`, which contains delivery records. 
Only completed deliveries which are goods-issued are taken into account. 

Table @tbl-02-data-sources-sales-filters shows the applied filters to select
standard sales orders and free-of-charge orders. 

::: {#tbl-02-data-sources-sales-filters .table}
```{r}

dtSF <-  fread(text = "
  Document Type, Item Category  , Reason
  TA           , TAN; ZTAQ; YGF1, standard sales orders
  KB; ZDR      ,                , standard sales orders
  TA           , TANN; ZTNG     , free-of-charge orders
  KL; ZFD      ,                , free-of-charge orders
  ")

# Create a styled kableExtra table
dtSF %>% 
  kbl(
    caption             = "Sales Data Filters",
    col.names           = names(.),
    align               = "l",
    booktabs            = TRUE
  ) %>% 
  kable_styling(
    full_width          = FALSE,
    position            = "left",
    bootstrap_options   = c("striped", "hover", "condensed"),
    latex_options       = c("striped", "hold_position")    
  )
```
Sales data filters
:::

::: {.callout-important title="Fixed Planned Promotions are excluded"}
Articles flagged in the Article Master data as Fixed Planned Promotions are excluded because they aren’t forecasted—instead, they’re delivered by contract and thus always achieve 100 % forecast accuracy.
:::

### Forecasts {#sec-02-forecasts}

Forecasts are generated monthly for each article–plant–customer combination, with a planning horizon of up to three years. For this research, the focus is on the year 2024, using a forecast horizon of up to 12 months. Consequently, only forecasts generated from 2023 onward are considered, see @fig-05-cv for visual illustration.

Since forecasts are created both before and after the Demand Review, the dataset is split into two versions: *pre-Demand Review* and *post-Demand Review*. The pre-Demand Review forecasts are used for the initial analysis, while the post-Demand Review forecasts serve as the basis for final validation.

## Data Collection Report {#sec-02-data-collection}
<!-- Summarize how data was obtained and any issues during acquisition: -->

Table @tbl-02-data-collection summarizes the data collection process, including the data sources, types, extraction dates, record counts, file sizes, and keys used for each data set. The data is collected from the SAP system and stored in a structured format for further processing.

Pre- and post-Demand Review forecasts are stored as separate files, one for each month in which they were created. Each file is named to reflect the extraction date.The key is  essential for linking the data sets during analysis.

::: {#tbl-02-data-collection .table}
```{r}

# Read the data into a data.table 
dtDC <-  fread(text = "
  Data Source, Type, Extract Date, Records, Size    , Files, Key
  article    , MD  , 21-01-2025  ,        , 1.691 KB,     3, Article     
  customer   , MD  , 21-01-2025  ,        , 1.345 KB,     2, Customer
  sales      , TD  , 21-01-2025  ,        , 1.7   GB,     4, Date; Article; Plant; Customer
  forecasts  , TD  , 21-01-2025  ,        , 8.99  GB,    48, Date; Article; Plant; CustHier3
")

# Create a styled kableExtra table
dtDC %>% 
  kbl(
    # caption             = "Data Collection ",
    col.names           = names(.),
    align               = "l",
    booktabs            = TRUE
  ) %>% 
  kable_styling(
    full_width          = FALSE,
    position            = "left",
    bootstrap_options   = c("striped", "hover", "condensed")
  )
```
Data Collection
:::


## Data Description Report {#sec-02-data-description}

<!-- Descriptive overview of structure and metadata: Open Hub definition -->
<!-- Variable types (date, categorical, numeric) -->
<!-- Dimensions (rows × columns) -->
<!-- Target variable (e.g., sales units, net revenue) -->

### Metadata {#sec-02-data-description-metadata}

Table @tbl-02-data-description-meta gives a sample of the datasets file name, extension, 
type of data and date of extraction. A detailed overview of all fields is provided in the appendix, @tbl-12-data-description-meta.

::: {#tbl-02-data-description-meta .table}
```{r}
# DT with Filter in Appendix
# dataLoads
 wb_read(
    file      = ET2BB,
    sheet     = "DATA_LOADS",
    start_row = 1,
    start_col = 1,
    skip_empty_cols = TRUE,
    skip_empty_rows = TRUE
  )  %T>%
  setDT() %>%
  .[RELEVANT == 'Y', !"RELEVANT"] %>%
  .[OHDEST %chin% c('DSCP_MATE', 'DSCP_CUST', 'OH_FRPR1')] %>%
  .[1:3] %>%
  .[, .(OHDEST, FILEEXTERN)] %>%  # , TYPE, DATE
  # .[is.na(KEY), KEY:= '']  %>%
  kbl(
    caption             = "Data Loads",
    col.names           = names(.),
    align               = "l",
    booktabs            = TRUE
  ) %>%
  kable_styling(
    full_width          = FALSE,
    position            = "left",
    bootstrap_options   = c("striped", "hover", "condensed")
  )
```
meta data on data extraction
:::

### Data fields {#sec-02-data-description-fields}

Table @tbl-02-data-description gives a sample of the dataset fields, including names, data types, key flags, conversion routines, and hierarchy levels. A detailed overview of all fields is provided in the appendix, @tbl-12-data-description.

::: {#tbl-02-data-description .table}
```{r}
# DT with Filter in Appendix
 wb_read(
    file      = ET2BB,
    sheet     = "DD",
    start_row = 1,
    start_col = 1,
    skip_empty_cols = TRUE,
    skip_empty_rows = TRUE
  )  %T>%
  setDT() %>%
  .[PYTHIA == 'Y', !"PYTHIA"] %>%
  .[c(1, 12, 33)] %>%
  kbl(
    caption             = "Data Description",
    col.names           = names(.),
    align               = "l",
    booktabs            = TRUE
  ) %>%
  kable_styling(
    full_width          = FALSE,
    position            = "left",
    bootstrap_options   = c("striped", "hover", "condensed")
  )
```
sample of overview of all data set fields.
:::


```{r}
#| label: tab-02-data-description

# Read the data into a data.table 
dtDD <-  fread(text = "
  Data Source, Date , Categorical, Numeric
  article    ,      ,         12 ,   0      
  customer   ,      ,         6  ,   0     
  sales      , Day  ,         7  ,   2     
  forecasts  , Month,         7  ,   2     
")

# Create a styled kableExtra table
dtDD %>% 
  kbl(
    caption             = "Data Description ",
    col.names           = names(.),
    align               = "l",
    booktabs            = TRUE
  ) %>% 
  kable_styling(
    full_width          = FALSE,
    position            = "left",
    bootstrap_options   = c("striped", "hover", "condensed")
  )
```

```{r}


ACC <-                              # 
  pa_td_dyn_get(
    .vtype       = c('010', '060')  , # 010 = Actuals, 060 = Forecast
    .ftype       = c(1, 2)          , # 1 = PreDR, 2 = PostDR 
    .material    = NULL      , # Optional user-supplied material
    .salesorg    = 'FR30'    , #'Optional user-supplied salesorg
    .scope_matl  = TRUE      , # restrict to Pythia Scope
    .scope_sorg  = TRUE      , # restrict to Pythia Scope
    .cm_min      = '202101'  , # minimal Cal Month
    .cm_max      = '202412'  , # maximal Cal Month
    .step_min    = 1         , # minimal forecast step ahead
    .step_max    = 18        , # maximal forecast step ahead
    .lagg_min    = -1        , # minimal diff. between VERSMON & MONTH
    .lagg_max    = -1        , # maximal diff. between VERSMON & MONTH  
    .n           = Inf  
  ) 



  wb_read(
    file      = ET2BB,
    sheet     = "ACC",
    start_row = 1,
    start_col = 1,
    skip_empty_cols = TRUE,
    skip_empty_rows = TRUE
  )  %>%
  setDT() %>%
  .[is.na(KEY), KEY:= ''] %>%
  kbl(
    caption             = "Sales & Forecast metadata",
    col.names           = names(.),
    align               = "l",
    booktabs            = TRUE
  ) %>% 
  kable_styling(
    full_width          = FALSE,
    position            = "left",
    bootstrap_options   = c("striped", "hover", "condensed")
  )


  wb_read(
    file      = ET2BB,
    sheet     = "ACC_TA",
    start_row = 1,
    start_col = 1,
    skip_empty_cols = TRUE,
    skip_empty_rows = TRUE
  )  %>% 
  kbl(
    caption             = "Sales & Forecast Fields, Unique Count and Example",
    col.names           = names(.),
    align               = "l",
    booktabs            = TRUE
  ) %>% 
  kable_styling(
    full_width          = FALSE,
    position            = "left",
    bootstrap_options   = c("striped", "hover", "condensed")
  )
```

## Data Exploration Report {#sec-02-data-exploration}

<!-- Exploratory analyses & sanity checks   -->
<!-- statistical properties and metadata. -->
<!-- Initial exploration to spot trends, seasonality, outliers, and segmentation patterns. -->
<!-- Subsections (as needed): -->
<!-- Sales coverage: distribution over time -->
<!-- ABC/XYZ, CV, zero-counts, sales length -->
<!-- Temporal patterns (weekly/monthly seasonality, trend, events) -->
<!-- Cross-sectional: contribution per customer/product/channel -->
<!-- exclude units pak and DS   -->

### Geographic Distribution of Sales {#sec-02-eda-geography}

::: {#fig-02-eda-geography .figure}

```{r}
#| fig-height: 4
#| fig-width: 6
#| 
dtMSO <- 
  wb_read(
    file      = ET2BB,
    sheet     = "SORG",
    start_row = 1,
    start_col = 1,
    skip_empty_cols = TRUE,
    skip_empty_rows = TRUE
  )                                   %T>%
  setDT()                              %>%
  setnames(
    c("SHARE"  , "BB-TEXT"),
    c("Percent", "MSO")
    )                                    %>%
  setorder(-Percent)                     %>%
  .[, Percent := Percent* 1]           %>%
  .[, ymax    := cumsum(Percent)]        %>%
  .[, ymin    := shift(ymax, fill = 0)]

# Sort factor levels so bars appear from smallest to largest (bottom to top)
dtMSO$MSO <- factor(dtMSO$MSO, levels = dtMSO$MSO[order(dtMSO$Percent)])

# Define a green color gradient: dark to light green
green_palette <- colorRampPalette(c("#38e56d", "#0f5e3c"))(7)  # darkgreen to lightgreen

# ggplot(dtMSO) +
#   geom_rect(aes(
#     xmin = 0.3, xmax = 0.7,
#     ymin = ymin, ymax = ymax,
#     fill = MSO
#   )) +
#   # coord_flip() +
#   scale_y_continuous(labels = scales::percent_format(scale = 1)) +
#   scale_fill_manual(values = green_palette) +
#   labs(
#     title = "Sales 2024 (in Value)",
#     # y     = NULL, # "Percentage",
#     x     = NULL,
#     fill  = "MSO"
#   ) +
#   theme_minimal() +
#   theme(
#     axis.text.x  = element_blank(), 
#     axis.text.y  = element_text(angle = 0),
#     axis.ticks.x = element_blank()
#   )

ggplot(dtMSO, aes(x = Percent, y = reorder(MSO, Percent))) +
  geom_col(fill = "forestgreen") +
  geom_text(aes(label = percent(Percent)), hjust = -0.1, size = 3.5) +
  scale_x_continuous(labels = percent_format(), limits = c(0, 0.6)) +
  labs(
    title = "Sales Share by Market Sales Organization (MSO)",
    x = "Share of Total Sales",
    y = NULL
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(hjust = 0, face = "bold"),
    plot.title.position = "plot",
    panel.grid.major.y = element_blank(),
    axis.text.y = element_text(face = "bold")
  )
```

Scoping the analysis to `r dtMSO[1, MSO]`, which accounts for 
`r paste(round( 100 * dtMSO[1, Percent], 2), '%')` of total sales in 2024.
:::

::: {.callout-note}
For modeling purposes, the data was scoped to A-class articles based on ABC classification.  
See [ABC Analysis and Scope Definition](#sec-03-abc) for details.
:::

### Data Profiling {#sec-02-data-profiling}

```{r}
#| label: skim-summary
#| results: 'asis'

dtACC_sk <- ACC %>%
  skim(.) %T>%
  setDT() %T>%
  setnames(c(
    "type",
    "variable", "n_missing", "complete_rate", 
    "min_length", "max_length", "empty", "n_unique", "whitespace", 
    "mean", "sd", "p0", "p25", "p50", "p75", "p100", "histogram"
  ))
```

#### Character Variables {#sec-02-data-profiling-ca}

Table @tbl-02-data-profiling-ca shows the data profiling results for the dataset,
including the number of missing values, completeness rate, minimum and maximum length 
for character variables. The profiling helps identify data quality issues and 
provides insights into the structure of the dataset.

::: {#tbl-02-data-profiling-ca .table}
```{r}
dtACC_sk[
  type == 'character', .(
    variable, 
    n_missing, complete_rate,
    min_length, max_length, empty, n_unique, whitespace
  )] %>% 
  kbl(
    # caption             = "Data Profiling: Character Variables",
    col.names           = names(.),
    align               = "l",
    booktabs            = TRUE
  ) %>% 
  kable_styling(
    full_width          = FALSE,
    position            = "left",
    bootstrap_options   = c("striped", "hover", "condensed")
  )
```
Data Profiling: Character Variables
:::

::: callout-note
This data set is of high quality and doesn't need any further analysis on 
the character variables.
:::


#### Numeric Variables {#sec-02-data-profiling-nu}

Table @tbl-02-data-profiling-nu shows the data profiling results for the 
numeric variables in the dataset, including mean, standard deviation, and 
percentiles. The histogram column provides a visual representation of 
the distribution of each numeric variable.

::: {#tbl-02-data-profiling-nu .table}
```{r}
dtACC_sk[
  type == 'numeric', .(
    variable, 
    mean, sd, p0, p25, p50, p75, p100, histogram
  )]  %>% 
  kbl(
    # caption             = "EDA: Numeric Variables",
    col.names           = names(.),
    align               = "l",
    booktabs            = TRUE
  ) %>% 
  kable_styling(
    full_width          = FALSE,
    position            = "left",
    bootstrap_options   = c("striped", "hover", "condensed")
  )
```
Data Profiling: Numeric Variables
:::

::: callout-important
We see that the Quantity (Q) needs some further analysis due to outliers.
:::


```{r}

dt <- ACC[MATERIAL != '000000000003020642' & VTYPE == '010']

ggplot(dt, aes(x = Q)) +
  # Histogram with density scaling, green fill and border
  geom_histogram(
    aes(y     = ..density..),
    binwidth  = diff(range(dt$Q)) / 100,  # adjust bin width (30 bins here)
    fill      = "lightgreen",
    color     = "forestgreen"
  ) +
  # Density curve on top
  geom_density(
    color     = "forestgreen",
    linewidth = 1
  ) +
  labs(
    title     = "Histogram of Q with Density Overlay",
    x         = "Q",
    y         = "Density"
  ) +
  theme_minimal()
```


<!-- } else { -->
<!--   dtMATL_sk %>% -->
<!--     kable("latex", booktabs = TRUE, caption = "Variable Description Report") %>% -->
<!--     kable_styling(latex_options = c("striped", "hold_position"), full_width = FALSE) -->
<!-- } -->

### Article Hierachy {#sec-02-data-description-hierarchy}

```{r}
#| label: tab-02-data-description-hierarchy

# Read the data into a data.table
dtMAT <- wb_read(
  file      = ET2BB,
  sheet     = "MASP",
  start_row = 19,
  start_col = 6,
  skip_empty_cols = TRUE,
  skip_empty_rows = TRUE
) %T>%
  setDT() %T>%
  setnames(
    c(4, 6, 8, 9), 
    c("ProdCat", "SubCat", "Family", "Article")
  ) %>%
  .[, .(ProdCat, SubCat, Family, Article)] %>%
  unique()

dtMAT_tab <- dtMAT[, .(
  SubCat_Count   = uniqueN(SubCat),
  Family_Count   = uniqueN(Family),
  Article_Count  = uniqueN(Article)
), by = ProdCat][order(ProdCat)]

dtCUST <- wb_read(
  file      = ET2BB,
  sheet     = "CUST",
  start_row = 19,
  start_col = 9,
  skip_empty_cols = TRUE,
  skip_empty_rows = TRUE
) %T>%
  setDT() %T>%
  setnames(c(1), c("KEY")  ) 

dtCUST_H1 <- wb_read(
  file      = ET2BB,
  sheet     = "CUST_H1",
  start_row = 1,
  start_col = 1,
  skip_empty_cols = TRUE,
  skip_empty_rows = TRUE
) %T>%
  setDT() %T>%
  setnames(c(1), c("KEY")  )

dtCST <- 
  dtCUST_H1[dtCUST, on = .(KEY), nomatch = 0] %T>%
  setnames(
    c(4, 7, 9, 11),
    c("Channel", "Key_Account", "Account", "SoldTo")
  ) %>%
  .[, .(Channel, Key_Account, Account, SoldTo, Customer)] %>%
  unique()

dtCST_tab <- dtCST[, .(
  KAM_count      = uniqueN(Key_Account),
  ACC_Count      = uniqueN(Account),
  SoldTo_Count   = uniqueN(SoldTo),
  Customer_Count = uniqueN(Customer)
), by = Channel][order(Channel)]
```

Table @tbl-02-data-description-hierarchy-article summarizes the 4-levels of the article hierarchy by a unique count per level. Roughly,

- each Family contains roughly `r mean(dtMAT_tab$Article_Count/dtMAT_tab$Family_Count) %>% round(0)` articles.
- each Sub Category contains roughly `r mean(dtMAT_tab$Family_Count/dtMAT_tab$SubCat_Count) %>% round(0)` families.

::: {#tbl-02-data-description-hierarchy-article .table}  
```{r}

datatable(
  dtMAT_tab,
    options      = list(
    dom          = 't',          # only the table ("t"); no filter, no pagination controls
    paging       = FALSE,        # disable pagination
    searching    = FALSE,        # disable search bar
    info         = FALSE,        # hide “Showing X of Y” info
    ordering     = TRUE,         # keep column sorting enabled (default = TRUE)
    lengthChange = FALSE,        # hide “Show N entries” dropdown
    columnDefs   = list(
      list(className = 'dt-left', targets = 0)
    )  
    )
,
  rownames     = FALSE,
  class        = 'stripe hover compact'
) 


```

Summary: Article Hierachy
:::

::: {.content-visible when-format="html"}

The following interactive chart allows to explore the hierarchy by 
clicking on the nodes to expand or collapse them. 
:::

:::: {#fig-inter_active_article_hier .figure width="15cm"}

::: {.content-visible when-format="html"}

```{r}
#| label: tab-02-data-description-hierarchy-plot

collapsibleTree(
  dtMAT,
  hierarchy   = c("ProdCat", "SubCat", "Family"),
  root        = "Articles",
  width       = "100%",    # now this 100% truly means “fill the parent”
  height      = "600px",   # you can adjust the height as needed
  zoomable    = TRUE,
  fillByLevel = TRUE,
  fill        = "#38e56d"
)
```
Interactive Article Hierarchy, click on the nodes to expand or collapse them.
:::

::::



### Customer Hierarchy {#sec-02-data-description-hierarchy-customer}

Table @tbl-02-data-description-hierarchy-customer summarizes the 4-levels of the 
customer hierarchy by a unique count per level. Roughly:

- each Family contains roughly 
`r mean(dtCST_tab$Customer_Count/dtCST_tab$SoldTo_Count, na.rm =TRUE) %>% round(0)` customers
- each Account contains roughly
`r mean(dtCST_tab$SoldTo_Count/dtCST_tab$ACC_Count, na.rm =TRUE) %>% round(0)` sold-to parties
- each Key Account contains roughly 
`r mean(dtCST_tab$ACC_Count/dtCST_tab$KAM_count, na.rm =TRUE) %>% round(0)` accounts,

::: {#tbl-02-data-description-hierarchy-customer .table}
```{r}

datatable(
  dtCST_tab,
  options = list(
    dom          = 't',          # only the table ("t"); no filter, no pagination controls
    paging       = FALSE,        # disable pagination
    searching    = FALSE,        # disable search bar
    info         = FALSE,        # hide “Showing X of Y” info
    ordering     = TRUE,         # keep column sorting enabled (default = TRUE)
    lengthChange = FALSE,       # hide “Show N entries” dropdown
    columnDefs   = list(
      list(className = 'dt-left', targets = 0)
    )
  ),
  rownames     = FALSE,
  class        = 'stripe hover compact'
) 

# dtCST_tab %>%
#   kbl(
#     caption             = "Summary: Customer Hierachy",
#     col.names           = names(.),
#     align               = "l",
#     booktabs            = TRUE
#   ) %>% 
#   kable_styling(
#     full_width          = FALSE,
#     position            = "left",
#     bootstrap_options   = c("striped", "hover", "condensed")
#   )

```

Summary: Customer Hierachy
:::

### Time Series Patterns {#sec-02-data-exploration-time-series}
let’s explore the time series patterns per category

the hot drinks category exhibits strong seasonality, as tea consumption increases during winter. 

Seasonality or trend plot
Using Time series features, categories can be automatically classified 
 
in the upper left corner, Clipper hot again, demonstrating the expected seasonality-strength.

### Forecastability {#sec-02-data-exploration-forecastability}

#### ABC-Classification {#sec-02-data-exploration-abc}
ABC/XYZ Classification
Purpose:
To prioritize items and assess whether forecasting is worthwhile or feasible for each.

Key Questions Addressed:

ABC: Is this item important enough to forecast or control tightly?
(based on contribution to revenue, volume, or margin)

XYZ: Is this item predictable enough to be reliably forecasted?
(based on variability, usually via Coefficient of Variation)

Typical Outcome:

High-priority & stable items (AX): Yes, forecast carefully.

Low-priority & erratic items (CZ): Maybe don't forecast — use reorder rules or make-to-order.

#### Syntetos-Boylan (SB) Classification {#sec-02-data-exploration-sb}

Purpose:
To guide model selection by identifying the nature of demand (e.g., intermittent, lumpy).

Key Questions Addressed:

What kind of statistical method is appropriate for this demand profile?

Is this demand pattern intermittent (zero-heavy), erratic, or smooth?

Model Implications:

Demand Type	Suggested Model
Smooth	Exponential smoothing (ETS)
Erratic	Damped trend, possibly ETS with adjustments
Intermittent	Croston (or variants: SBA, TSB, etc.)
Lumpy	Advanced models (e.g., bootstrapped, hybrid, or judgmental overlays)

Smooth: Low ADI and low CV. Highly predictable demand patterns.
 • Intermittent: High ADI, low CV. Forecastable with specialized methods such as
 Croston’s, ADIDA, or IMAPA.
 • Erratic: Low ADI, high CV. High uncertainty, often driven by price sensitivity or
 promotions.
 • Lumpy: High ADI and high CV. Extremely difficult to forecast, often related to slow
moving or event-driven items.

```{r}

fread(text = "
SB_Type,Best_fit_Methods
Smooth,SES; Holt’s; Holt-Winters (Triple Exp); ARIMA; Moving Average
Erratic,Damped exponential; Holt-Winters (if seasonal); ARIMA with volatility modeling
Intermittent,Croston; SBA; TSB; bootstrapped quantile models
Lumpy,Croston variants + heuristics; simulation; judgmental overlays
") %>%
  kbl(
    # caption   = "Forecast Use Cases by Department",
    col.names = names(.),
    align     = "l",
    booktabs  = TRUE
  ) %>%
  kable_styling(
    full_width = FALSE,
    position   = "left",
    bootstrap_options = c("striped", "hover", "condensed")
  )

```


```{r}
fread(text = "
Method,Assumes,Handles,Fit_in_SB_Classification,Notes
Moving Average,Stationarity; no trend or seasonality,Smooth to moderate variability,\"Smooth; maybe Erratic\",Simple; not suitable for intermittent or seasonal
Simple Exponential Smoothing (SES),Stationary; no trend or seasonality,Random variation; no trend/season,\"Smooth\",Cannot handle trend or seasonality
Double Exponential Smoothing,Linear trend,Trend,\"Smooth; Erratic\",Better than SES for trending series
Triple Exponential Smoothing (Holt-Winters),Trend + seasonality,Seasonality,\"Smooth; some Erratic\",Not appropriate for intermittent/lumpy
ARIMA / ARMA,Stationarity (or with differencing),Trend + autocorrelation,\"Smooth; Erratic\",Requires regular frequency and sufficient data
Croston / SBA / TSB,Zero-heavy demand,Intermittency,\"Intermittent; Lumpy\",Designed specifically for intermittent demand
")%>%
  kbl(
    # caption   = "Forecast Use Cases by Department",
    col.names = names(.),
    align     = "l",
    booktabs  = TRUE
  ) %>%
  kable_styling(
    full_width = FALSE,
    position   = "left",
    bootstrap_options = c("striped", "hover", "condensed")
  )

```

```{r}
library(data.table)

fread(text = "
Aspect,XYZ,ADI_CV2_Syntetos_Boylan
Metric,Coefficient of Variation (CV),ADI + CV² (more granular)
Forecastability,Measures volatility,Measures both volatility and intermittency
Use case,Broad segmentation for policy setting,Model selection and demand type understanding
Z class equivalent,Often aligns with 'Lumpy' or 'Erratic',Differentiates between those
") %>%
  kbl(
    # caption   = "Forecast Use Cases by Department",
    col.names = names(.),
    align     = "l",
    booktabs  = TRUE
  ) %>%
  kable_styling(
    full_width = FALSE,
    position   = "left",
    bootstrap_options = c("striped", "hover", "condensed")
  )

```


## Data Quality Report {#sec-02-data-quality}

<!-- Assess data integrity and readiness for modeling. -->
<!-- Subsections: -->
<!-- Missing values (which variables, how much, patterns) -->
<!-- Consistency checks (e.g., units vs revenue) -->
<!-- Duplicated records -->
<!-- Outliers: any anomalous spikes or drops? -->

### Data Quality Assessment {#sec-02-data-quality-assessment}
<!-- Assessment: What issues were found (missing values, outliers, etc.) -->

<!-- Data Quality Assessment and Remediation -->
<!-- Summarize actions taken: -->
<!-- Imputation strategies -->
<!-- Exclusions (e.g., rare SKUs or obsolete accounts) -->

all 0 
duplicates, zero quantity, unexpected NAs

**Actions taken:** 
<!-- Any exclusions or imputations -->
<!-- Summarize actions taken: -->

## Summary of Data Understanding {#sec-02-summary}

- The dataset combines master and transactional data from SAP BI and QAD systems.
- Sales and forecasts are scoped to a single MSO and A-class finished goods.
- Product and customer hierarchies follow a consistent four-level structure.
- Exploratory analysis confirms sufficient data volume, patterns, and quality.
- The data is now ready for scoping, transformation, and modeling.


<!-- {{< include 02_data_understanding_def.qmd >}} -->
