---
subtitle:     what data is available and where it comes from 
description: |
  - SAP BI extracts serve as the foundation: article, customer, sales & forecast data
  - Data was scoped to finished goods under Make-to-Stock with ambient storage
  - The modeling dataset is based on the top-performing MSO and A-class articles
  - Exploratory analysis confirmed temporal seasonality and product hierarchy depth
  - Data quality is high overall, with minimal missing values and no major outliers
number-sections: false
toc: true         # show page‐TOC (right‐sidebar) on _every_ page
toc-depth: 3      # include ## and ### in the TOC
---

# Data Understanding {#sec-02-data-understanding}

```{r}
#| label: setup
source("../_book_functions.R")
```

<!-- ## Identification of Data Sources -->


<!-- ### Articles -->

<!-- ### Customers -->


<!-- ### Sales -->
<!-- Data filters applied (Doctype's, Item category, ...) -->
<!-- exclude Promotion articles --> 

<!-- ### Forecasts -->

<!-- ## Data Collection Report -->
<!-- Summarize how data was obtained and any issues during acquisition: -->


## Data Description Report {#sec-12-data-description}

<!-- Descriptive overview of structure and metadata: Open Hub definition -->
<!-- Variable types (date, categorical, numeric) -->
<!-- Dimensions (rows × columns) -->
<!-- Target variable (e.g., sales units, net revenue) -->

### Metadata Overview {#sec-12-data-description-metadata}

Table @tbl-02-data-description-meta gives all details of the datasets,
back sec-02-data-description-metadata 

::: {tbl-12-data-description-meta .table}
```{r}
# DT with Filter in Appendix
# dataLoads
 wb_read(
    file      = ET2BB,
    sheet     = "DATA_LOADS",
    start_row = 1,
    start_col = 1,
    skip_empty_cols = TRUE,
    skip_empty_rows = TRUE
  )  %T>%
  setDT() %>%
  .[RELEVANT == 'Y', !"RELEVANT"] %>%
  # .[OHDEST %chin% c('DSCP_MATE', 'DSCP_CUST', 'OH_FRPR1')] %>%
  # .[, .(OHDEST, FILEEXTERN, TYPE, DATE)]
  # .[is.na(KEY), KEY:= ''] %>%
  kbl(
    caption             = "Data Loads",
    col.names           = names(.),
    align               = "l",
    booktabs            = TRUE
  ) %>%
  kable_styling(
    full_width          = FALSE,
    position            = "left",
    bootstrap_options   = c("striped", "hover", "condensed")
  )
```
meta data on data extraction
:::

### Data fields {#sec-12-data-description-fields}

A detailed overview of all dataset fields, including names, data types, key flags, conversion routines, and hierarchy levels, back to @sec-02-data-description

::: {#tbl-12-data-description .table}
```{r}
# DT with Filter in Appendix
 wb_read(
    file      = ET2BB,
    sheet     = "DD",
    start_row = 1,
    start_col = 1,
    skip_empty_cols = TRUE,
    skip_empty_rows = TRUE
  )  %T>%
  setDT() %>%
  .[PYTHIA == 'Y', !"PYTHIA"] %>%
  # .[c(1, 12, 23)] %>%
  kbl(
    caption             = "Data Description",
    col.names           = names(.),
    align               = "l",
    booktabs            = TRUE
  ) %>%
  kable_styling(
    full_width          = FALSE,
    position            = "left",
    bootstrap_options   = c("striped", "hover", "condensed")
  )
```
A detailed overview of all dataset fields.
:::

<!-- ## Data Exploration Report -->

<!-- Exploratory analyses & sanity checks   -->
<!-- statistical properties and metadata. -->
<!-- Initial exploration to spot trends, seasonality, outliers, and segmentation patterns. -->
<!-- Subsections (as needed): -->
<!-- Sales coverage: distribution over time -->
<!-- ABC/XYZ, CV, zero-counts, sales length -->
<!-- Temporal patterns (weekly/monthly seasonality, trend, events) -->
<!-- Cross-sectional: contribution per customer/product/channel -->
<!-- exclude units pak and DS   -->

<!-- ### Geographic Distribution of Sales -->


### Data Profiling

```{r}
#| label: skim-summary
#| results: 'asis'

ACC <-                              # 
  pa_td_dyn_get(
    .vtype       = c('010', '060')  , # 010 = Actuals, 060 = Forecast
    .ftype       = c(1, 2)          , # 1 = PreDR, 2 = PostDR 
    .material    = NULL      , # Optional user-supplied material
    .salesorg    = 'FR30'    , #'Optional user-supplied salesorg
    .scope_matl  = TRUE      , # restrict to Pythia Scope
    .scope_sorg  = TRUE      , # restrict to Pythia Scope
    .cm_min      = '202101'  , # minimal Cal Month
    .cm_max      = '202412'  , # maximal Cal Month
    .step_min    = 1         , # minimal forecast step ahead
    .step_max    = 18        , # maximal forecast step ahead
    .lagg_min    = -1        , # minimal diff. between VERSMON & MONTH
    .lagg_max    = -1        , # maximal diff. between VERSMON & MONTH  
    .n           = Inf  
  ) 



dtACC_sk <- ACC %>%
  skim(.) %T>%
  setDT() %T>%
  setnames(c(
    "type",
    "variable", "n_missing", "complete_rate", 
    "min_length", "max_length", "empty", "n_unique", "whitespace", 
    "mean", "sd", "p0", "p25", "p50", "p75", "p100", "histogram"
  ))
```

#### Character Variables

Table @tbl-02-data-profiling-ca shows the data profiling results for the dataset,
including the number of missing values, completeness rate, minimum and maximum length 
for character variables. The profiling helps identify data quality issues and 
provides insights into the structure of the dataset.

::: {#tbl-02-data-profiling-ca}
```{r}
dtACC_sk[
  type == 'character', .(
    variable, 
    n_missing, complete_rate,
    min_length, max_length, empty, n_unique, whitespace
  )] %>% 
  kbl(
    # caption             = "Data Profiling: Character Variables",
    col.names           = names(.),
    align               = "l",
    booktabs            = TRUE
  ) %>% 
  kable_styling(
    full_width          = FALSE,
    position            = "left",
    bootstrap_options   = c("striped", "hover", "condensed")
  )
```
Data Profiling: Character Variables
:::

::: callout-note
This data set is of high quality and doesn't need any further analysis on 
the character variables.
:::


#### Numeric Variables

Table @tbl-02-data-profiling-nu shows the data profiling results for the 
numeric variables in the dataset, including mean, standard deviation, and 
percentiles. The histogram column provides a visual representation of 
the distribution of each numeric variable.

::: {#tbl-02-data-profiling-nu}
```{r}
dtACC_sk[
  type == 'numeric', .(
    variable, 
    mean, sd, p0, p25, p50, p75, p100, histogram
  )]  %>% 
  kbl(
    # caption             = "EDA: Numeric Variables",
    col.names           = names(.),
    align               = "l",
    booktabs            = TRUE
  ) %>% 
  kable_styling(
    full_width          = FALSE,
    position            = "left",
    bootstrap_options   = c("striped", "hover", "condensed")
  )
```
Data Profiling: Numeric Variables
:::

::: callout-important
We see that the Quantity (Q) needs some further analysis due to outliers.
:::


```{r}

dt <- ACC[MATERIAL != '000000000003020642' & VTYPE == '010']

ggplot(dt, aes(x = Q)) +
  # Histogram with density scaling, green fill and border
  geom_histogram(
    aes(y     = ..density..),
    binwidth  = diff(range(dt$Q)) / 100,  # adjust bin width (30 bins here)
    fill      = "lightgreen",
    color     = "forestgreen"
  ) +
  # Density curve on top
  geom_density(
    color     = "forestgreen",
    linewidth = 1
  ) +
  labs(
    title     = "Histogram of Q with Density Overlay",
    x         = "Q",
    y         = "Density"
  ) +
  theme_minimal()
```


<!-- } else { -->
<!--   dtMATL_sk %>% -->
<!--     kable("latex", booktabs = TRUE, caption = "Variable Description Report") %>% -->
<!--     kable_styling(latex_options = c("striped", "hold_position"), full_width = FALSE) -->
<!-- } -->

### Article Hierachy

```{r}
#| label: tab-02-data-description-hierarchy

# Read the data into a data.table
dtMAT <- wb_read(
  file      = ET2BB,
  sheet     = "MASP",
  start_row = 19,
  start_col = 6,
  skip_empty_cols = TRUE,
  skip_empty_rows = TRUE
) %T>%
  setDT() %T>%
  setnames(
    c(4, 6, 8, 9), 
    c("ProdCat", "SubCat", "Family", "Article")
  ) %>%
  .[, .(ProdCat, SubCat, Family, Article)] %>%
  unique()

dtMAT_tab <- dtMAT[, .(
  SubCat_Count   = uniqueN(SubCat),
  Family_Count   = uniqueN(Family),
  Article_Count = uniqueN(Article)
), by = ProdCat][order(ProdCat)]

dtCUST <- wb_read(
  file      = ET2BB,
  sheet     = "CUST",
  start_row = 19,
  start_col = 9,
  skip_empty_cols = TRUE,
  skip_empty_rows = TRUE
) %T>%
  setDT() %T>%
  setnames(c(1), c("KEY")  ) 

dtCUST_H1 <- wb_read(
  file      = ET2BB,
  sheet     = "CUST_H1",
  start_row = 1,
  start_col = 1,
  skip_empty_cols = TRUE,
  skip_empty_rows = TRUE
) %T>%
  setDT() %T>%
  setnames(c(1), c("KEY")  )

dtCST <- 
  dtCUST_H1[dtCUST, on = .(KEY), nomatch = 0] %T>%
  setnames(
    c(4, 7, 9, 11),
    c("Channel", "Key_Account", "Account", "SoldTo")
  ) %>%
  .[, .(Channel, Key_Account, Account, SoldTo, Customer)] %>%
  unique()

dtCST_tab <- dtCST[, .(
  KAM_count      = uniqueN(Key_Account),
  ACC_Count      = uniqueN(Account),
  SoldTo_Count   = uniqueN(SoldTo),
  Customer_Count = uniqueN(Customer)
), by = Channel][order(Channel)]
```

Table @tbl-02-data-description-hierarchy-article summarizes the 4-levels of the article hierarchy by a unique count per level. Roughly,

- each Family contains roughly `r mean(dtMAT_tab$Article_Count/dtMAT_tab$Family_Count) %>% round(0)` articles.
- each Sub Category contains roughly `r mean(dtMAT_tab$Family_Count/dtMAT_tab$SubCat_Count) %>% round(0)` families.

::: {#tbl-02-data-description-hierarchy-article}  
```{r}

datatable(
  dtMAT_tab,
    options      = list(
    dom          = 't',          # only the table ("t"); no filter, no pagination controls
    paging       = FALSE,        # disable pagination
    searching    = FALSE,        # disable search bar
    info         = FALSE,        # hide “Showing X of Y” info
    ordering     = TRUE,         # keep column sorting enabled (default = TRUE)
    lengthChange = FALSE,        # hide “Show N entries” dropdown
    columnDefs   = list(
      list(className = 'dt-left', targets = 0)
    )  
    )
,
  rownames     = FALSE,
  class        = 'stripe hover compact'
) 


```

Summary: Article Hierachy
:::

::: {.content-visible when-format="html"}

The following interactive chart allows to explore the hierarchy by 
clicking on the nodes to expand or collapse them. 
:::

:::: {#fig-inter_active_article_hier .figure width="15cm"}

::: {.content-visible when-format="html"}

```{r}
#| label: tab-02-data-description-hierarchy-plot

collapsibleTree(
  dtMAT,
  hierarchy   = c("ProdCat", "SubCat", "Family"),
  root        = "Articles",
  width       = "100%",    # now this 100% truly means “fill the parent”
  height      = "600px",   # you can adjust the height as needed
  zoomable    = TRUE,
  fillByLevel = TRUE,
  fill        = "#38e56d"
)
```
Interactive Article Hierarchy, click on the nodes to expand or collapse them.
:::

::::



### Customer Hierarchy

Table @tbl-02-data-description-hierarchy-customer summarizes the 4-levels of the 
customer hierarchy by a unique count per level. Roughly:

- each Family contains roughly 
`r mean(dtCST_tab$Customer_Count/dtCST_tab$SoldTo_Count, na.rm =TRUE) %>% round(0)` customers
- each Account contains roughly
`r mean(dtCST_tab$SoldTo_Count/dtCST_tab$ACC_Count, na.rm =TRUE) %>% round(0)` sold-to parties
- each Key Account contains roughly 
`r mean(dtCST_tab$ACC_Count/dtCST_tab$KAM_count, na.rm =TRUE) %>% round(0)` accounts,

::: {#tbl-02-data-description-hierarchy-customer}
```{r}

datatable(
  dtCST_tab,
  options = list(
    dom          = 't',          # only the table ("t"); no filter, no pagination controls
    paging       = FALSE,        # disable pagination
    searching    = FALSE,        # disable search bar
    info         = FALSE,        # hide “Showing X of Y” info
    ordering     = TRUE,         # keep column sorting enabled (default = TRUE)
    lengthChange = FALSE,       # hide “Show N entries” dropdown
    columnDefs   = list(
      list(className = 'dt-left', targets = 0)
    )
  ),
  rownames     = FALSE,
  class        = 'stripe hover compact'
) 

# dtCST_tab %>%
#   kbl(
#     caption             = "Summary: Customer Hierachy",
#     col.names           = names(.),
#     align               = "l",
#     booktabs            = TRUE
#   ) %>% 
#   kable_styling(
#     full_width          = FALSE,
#     position            = "left",
#     bootstrap_options   = c("striped", "hover", "condensed")
#   )

```

Summary: Customer Hierachy
:::

<!-- ### Time Series Patterns -->


## Data Quality Report

<!-- Assess data integrity and readiness for modeling. -->
<!-- Subsections: -->
<!-- Missing values (which variables, how much, patterns) -->
<!-- Consistency checks (e.g., units vs revenue) -->
<!-- Duplicated records -->
<!-- Outliers: any anomalous spikes or drops? -->

## Data Quality Assessment

<!-- Data Quality Assessment and Remediation -->
<!-- Summarize actions taken: -->
<!-- Imputation strategies -->
<!-- Exclusions (e.g., rare SKUs or obsolete accounts) -->

all 0 
duplicates, zero quantity, unexpected NAs

## Summary of Data Understanding {#sec-02-summary}

- The dataset combines master and transactional data from SAP BI and QAD systems.
- Sales and forecasts are scoped to a single MSO and A-class finished goods.
- Product and customer hierarchies follow a consistent four-level structure.
- Exploratory analysis confirms sufficient data volume, patterns, and quality.
- The data is now ready for scoping, transformation, and modeling.


<!-- {{< include 02_data_understanding_def.qmd >}} -->
