---
subtitle:    Beat the Nerd! Forecast smart-R with Augmented Insights at Scale
description: |
  - *Pythia’s Advice* is live and integrated with SAP BW and SAC.
  - Monthly refreshes provide updated forecasts and model insights.
  - "Beat the Nerd" gamification builds trust and boosts accuracy.
  - Docker, Kubernetes, and Git enable scalable, maintainable deployment.
  - ETS prediction intervals support safety stock and reduce waste.
  - Monitoring and feedback loops keep the system aligned with needs.
number-sections: false
toc: true         # show page‐TOC (right‐sidebar) on _every_ page
toc-depth: 3      # include ## and ### in the TOC
---

# Deployment {#sec-06-deployment}

```{r}
#| label: setup
#| echo: false
#| include: false
source("../_book_functions.R")
```

```{r}
#| label: Features * Classifaction data data
#| echo: false
#| include: false
#| message: false
#| warning: false
#| results: hide
#| eval: true

dtAUG <- readRDS(file.path(PDATA, "dtAUG.rds")) %>%
  .[, `:=`(
    M_value = RL0(M_value),
    C_value = RL0(C_value)
  )    ]

```

<!-- ## Overview {#sec-16-deployment-overview} -->


## Operationalized Proof-of-Concepts {#sec-16-deployment-proof-of-concepts}


### Statistical Indicators {#sec-06-deployment-statistical-indicators}

The *Pythia’s Advice* system generates key statistical indicators for each time series, including:

-   **SLEN**: Represents the length of the time series, indicating how many historical data points are available for forecasting. This helps planners assess the reliability of the forecast.
-   **minCM**: Indicates the first Month of the time series, providing context for the historical data and its relevance to current demand patterns.
-   **maxCM**: Represents the last Month of the time series, helping planners understand the recency of the data and its implications for forecasting.
-   **N**: The number of observations in the time series, which is crucial for determining the Average Demand Interval (ADI).
-   **Q**: Total Quantity, which is the sum of all historical demand values. This metric helps planners understand the overall demand volume for the time series.
-   **MU**: The Mean of the time series, providing a central tendency measure that helps planners gauge typical demand levels.
-   **SIGMA**: The Standard Deviation, which measures the variability of demand. This helps planners assess the stability of the time series.


::: {#tbl-16-deployment-statistical-indicators} 

```{r}

datatable(
  dtAUG[sample(1:.N, 100),
        .(combination, M = M_value, C = C_value, 
          minCM      , maxCM  , SLEN   , 
          N          , Q      ,
          MU    = round(MU    , 2)   , 
          SIGMA = round(SIGMA , 2)   , 
          ADI   = round(ADI   , 2)   ,
          CV    =  round(CV   , 2)   ,
          CV2   = round(CV2   , 2))],
  options = list(
    paging     = TRUE,        #  pagination
    searching  = TRUE,        # disable search bar
    info       = TRUE,        # hide “Showing X of Y” info
    ordering   = TRUE,         # keep column sorting enabled (default = TRUE)    
    pageLength = 10,
    lengthMenu = c(5, 10, 25, 50, 100, 2000),
    autoWidth  = TRUE,
    scrollX    = TRUE,
    columnDefs   = list(
      list(className = 'dt-left', targets = 0)
    )
    # , 
    # searchCols = list(
    #   NULL,               # No filter on column OHDEST
    #   NULL,               # No filter on column DATATYPE
    #   NULL,               # No filter on column FLDNM_IN
    #   NULL,               # No filter on column FIELDTP      
    #   NULL,               # No filter on column TRNSFRM
    #   NULL,               # No filter on column FLDNM_OUT      
    #   list(search = "C")  # Filter column SRC on value "C"
    #   # Add more elements for other columns if needed
    # )
  ),
  filter     = "top",  # or "bottom",  
  rownames     = FALSE,
  class        = 'stripe hover compact'
) %>%
  formatStyle(
    columns    = "combination",  # Target just the first column
    fontWeight = "bold"          # Make it bold
  ) %>%
  formatStyle(
    columns  = c("combination",
      "M"      , "C"      , "minCM", "maxCM", "SLEN"   , 
      "N"      , "Q"      , "MU"   , 
      "SIGMA"  , "ADI"    ,
      "CV"     ,  "CV2"  
      ),
    fontSize = '0.8em'          # reduce font size for other columns if needed
  )


          
```

**Interactive Table**: Statistical indicators per Aggregation Level.
:::


### Classifications {#sec-16-deployment-classifications}

The *Pythia’s Advice* system generates comprehensive time series classifications, including:  

-   **ABC Classification**: Prioritizes time series based on their contribution to overall demand, allowing planners to focus on the most impactful items.

-   **XYZ Classification**: Assesses the variability of demand patterns, helping planners identify stable versus erratic time series.

-   **SBC Classification**: Evaluates the seasonality of time series, enabling planners to adjust forecasts based on seasonal trends.


::: {#tbl-16-deployment-classification}  
```{r}

datatable(
  dtAUG[sample(1:.N, 100), 
        .(combination, M = M_value, C = C_value, ABC, XYZ, SBC)],
  options = list(
    paging     = TRUE,        #  pagination
    searching  = TRUE,        # disable search bar
    info       = TRUE,        # hide “Showing X of Y” info
    ordering   = TRUE,         # keep column sorting enabled (default = TRUE)    
    pageLength = 10,
    lengthMenu = c(5, 10, 25, 50, 100, 2000),
    autoWidth  = TRUE,
    scrollX    = TRUE,
    columnDefs   = list(
      list(className = 'dt-left', targets = 0)
    )
    # , 
    # searchCols = list(
    #   NULL,               # No filter on column OHDEST
    #   NULL,               # No filter on column DATATYPE
    #   NULL,               # No filter on column FLDNM_IN
    #   NULL,               # No filter on column FIELDTP      
    #   NULL,               # No filter on column TRNSFRM
    #   NULL,               # No filter on column FLDNM_OUT      
    #   list(search = "C")  # Filter column SRC on value "C"
    #   # Add more elements for other columns if needed
    # )
  ),
  filter     = "top",  # or "bottom",  
  rownames     = FALSE,
  class        = 'stripe hover compact'
) %>%
  formatStyle(
    columns    = "combination",  # Target just the first column
    fontWeight = "bold"          # Make it bold
  ) %>%
  formatStyle(
    columns  = c("combination", "M", "C", "ABC", "XYZ", "SBC"),
    fontSize = '0.8em'          # reduce font size for other columns if needed
  )

```

**Interactive Table**: Classifications per Aggregation Level
:::

### Seasonal & Trend Indicators {#sec-06-deployment-seasonal-trend}

The *Pythia’s Advice* system also provides insights into seasonal and trend components of the time series:
-   **Seasonal**: These metrics help planners understand the seasonal patterns in demand, enabling them to adjust forecasts based on expected seasonal fluctuations.
-   **Trend**: These metrics capture the underlying trends in demand over time, allowing planners to identify long-term shifts in demand patterns.
-   **Spikiness**: Measures the degree of irregular spikes in demand, helping planners identify potential anomalies or outliers.
-   **Linearity**: Assesses the linearity of the demand pattern, indicating how consistent the demand is over time.

::: {#tbl-16-deployment-seasonal-trend}
```{r}
datatable(
dtAUG[sample(1:.N, 100), 
      .(combination, M = M_value, C = C_value, 
        trend      = round(trend_strength      , 2), 
        Season     = round(seasonal_strength_12, 2),
        Peak       = round(seasonal_peak_12    , 2), 
        Through    = round(seasonal_trough_12  , 2),
        Spikeiness = round(spikiness           , 2), 
        Linearity  = round(linearity           , 2)
        )],
  options = list(
    paging     = TRUE,        #  pagination
    searching  = TRUE,        # disable search bar
    info       = TRUE,        # hide “Showing X of Y” info
    ordering   = TRUE,         # keep column sorting enabled (default = TRUE)    
    pageLength = 10,
    lengthMenu = c(5, 10, 25, 50, 100, 2000),
    autoWidth  = TRUE,
    scrollX    = TRUE,
    columnDefs   = list(
      list(className = 'dt-left', targets = 0)
    )
    # , 
    # searchCols = list(
    #   NULL,               # No filter on column OHDEST
    #   NULL,               # No filter on column DATATYPE
    #   NULL,               # No filter on column FLDNM_IN
    #   NULL,               # No filter on column FIELDTP      
    #   NULL,               # No filter on column TRNSFRM
    #   NULL,               # No filter on column FLDNM_OUT      
    #   list(search = "C")  # Filter column SRC on value "C"
    #   # Add more elements for other columns if needed
    # )
  ),
  filter     = "top",  # or "bottom",  
  rownames     = FALSE,
  class        = 'stripe hover compact'
) %>%
  formatStyle(
    columns    = "combination",  # Target just the first column
    fontWeight = "bold"          # Make it bold
  ) %>%
  formatStyle(
    columns  = c("combination", 
      "M", "C", "trend", "Season", "Peak", 
      "Through", "Spikeiness", "Linearity"),
    fontSize = '0.8em'          # reduce font size for other columns if needed
  )

```

**Interactive Table**: Seasonal and Trend Indicators per Aggregation Level

:::

<!-- ## Gamification {#sec-16-deployment-gamification} -->


<!-- ## Prediction Intervals {#sec-16-deployment-prediction-intervals} -->



<!-- ## Model Management {#sec-16-deployment-model-management} -->

<!-- ### Scalability {#sec-16-deployment-scalability} -->


<!-- ### Model Versioning {#sec-16-deployment-versioning} -->


<!-- ### Maintenance & Updating the model {#sec-16-deployment-maintenance} -->

<!-- ### Monitoring the Model {#sec-16-deployment-monitoring} -->


<!-- ## Summary {#sec-16-deployment-summary} -->


<!-- {{< include 16_deployment_def.qmd >}} -->
